---
title: "Day 4 Interactive Session"
author: "Alex Vand"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

## Code warm-up

You run two miles, and these are your quarter mile splits (in minutes). Write a for loop that returns your total time for any two consecutive quarter miles.

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)
half_splits <- vector(mode = "numeric", length = length(quarter_splits) - 1)

for(i in 1:(length(quarter_splits) - 1)) {
  half_splits[i] <- quarter_splits[i] + quarter_splits[i + 1]
}

half_splits
```

## Basic function structure practice

```{r}
birddog_sum <- function(bird, dog) {
  
  if(bird > 50) {
    stop("Birds cannot exceed 50.")
  }
  
  if(bird < 10) {
    message("Cool little bird!")
  }
  
  if(dog < 0) {
    warning("Can dog value be negative?")
  }
  
  add_birddog <- bird + dog
  return(add_birddog)
}


bird100_dog5 <- birddog_sum(bird = 7, dog = -5)
bird100_dog5

# birddog_sum(bird = "hawk", dog = 10) # error due to string input
```

```{r}
say_hello <- function(friend, weekday) {
  print(paste("Hello,", friend, "have a great", weekday))
}

say_hello(friend = "Teddy", weekday = "Thursday")

```

### Print vs. return example

```{r}
find_multiple <- function(a,b) {
  my_multiplication <- a * b
  print(my_multiplication)
}

find_multiple(a = 10, b = 6) + 12
# in py, this would return an error message
```

### Functions with conditions

```{r}
animal_age <- function(animal, age) {
  
  if(!animal %in% c("dog", "goat")) {
    stop("Animal must be a dog or a goat.")
  }
  
  if(animal == "dog") {
    creature_age <- age * 7
  }
  
  else if(animal == "goat") {
    creature_age <- age * 4.7
  }
  
  return(creature_age)
}

animal_age(animal = "goat", age = 13)
# animal_age(animal = "zebra", age = 66)
```

### Using the in operator

```{r}
adelie_gentoo <- penguins %>% 
  filter(species %in% c("Adelie", "Gentoo"))
```


Wrong example:

```{r}
adelie_gentoo_wrong <- penguins %>% 
  filter(species == c("Adelie", "Gentoo"))
```


### Functions with for loops

```{r}
df_means <- function(df) {
  for(i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    print(col_mean)
  }
}
# colnames(df[i]) will return that column name as a string

df_means(df = mtcars)

# Update this function so that for each iteration it returns a phrase "The mean value of _____ is _____."

df_means <- function(df) {
  for(i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    print(paste("The mean value of", colnames(df[i]), "is", col_mean))
  }
}
```


Make a subset of only structural size variables from the penguins dataset for Chinstrap

```{r}
chinstrap_sizes <- penguins %>% 
  filter(species == "Chinstrap") %>% 
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)

df_means(df = chinstrap_sizes)
```

## Condition to check class of an input

```{r}
lunchtime <- function(tacos, price_per) {
  
  if(is.numeric(tacos) == FALSE) {
    stop("You must input a number of tacos.")
  }
  
  tot_price <- tacos * price_per
  return(tot_price)
}

lunchtime(tacos = 3, price_per = 2.50)
# lunchtime(tacos = "chicken", price_per = 2.50)
```












