---
title: "Day 4 Interactive Session"
author: "Alex Vand"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

## Code warm-up

You run two miles, and these are your quarter mile splits (in minutes). Write a for loop that returns your total time for any two consecutive quarter miles.

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)
half_splits <- vector(mode = "numeric", length = length(quarter_splits) - 1)

for(i in 1:(length(quarter_splits) - 1)) {
  half_splits[i] <- quarter_splits[i] + quarter_splits[i + 1]
}

half_splits
```

## Basic function structure practice

```{r}
birddog_sum <- function(bird, dog) {
  
  if(bird > 50) {
    stop("Birds cannot exceed 50.")
  }
  
  if(bird < 10) {
    message("Cool little bird!")
  }
  
  if(dog < 0) {
    warning("Can dog value be negative?")
  }
  
  add_birddog <- bird + dog
  return(add_birddog)
}


bird100_dog5 <- birddog_sum(bird = 7, dog = -5)
bird100_dog5

# birddog_sum(bird = "hawk", dog = 10) # error due to string input
```

```{r}
say_hello <- function(friend, weekday) {
  print(paste("Hello,", friend, "have a great", weekday))
}

say_hello(friend = "Teddy", weekday = "Thursday")

```

### Print vs. return example

```{r}
find_multiple <- function(a,b) {
  my_multiplication <- a * b
  print(my_multiplication)
}

find_multiple(a = 10, b = 6) + 12
# in py, this would return an error message
```

### Functions with conditions

```{r}
animal_age <- function(animal, age) {
  
  if(!animal %in% c("dog", "goat")) {
    stop("Animal must be a dog or a goat.")
  }
  
  if(animal == "dog") {
    creature_age <- age * 7
  }
  
  else if(animal == "goat") {
    creature_age <- age * 4.7
  }
  
  return(creature_age)
}

animal_age(animal = "goat", age = 13)
# animal_age(animal = "zebra", age = 66)
```

### Using the in operator

```{r}
adelie_gentoo <- penguins %>% 
  filter(species %in% c("Adelie", "Gentoo"))
```


Wrong example:

```{r}
adelie_gentoo_wrong <- penguins %>% 
  filter(species == c("Adelie", "Gentoo"))
```


### Functions with for loops

```{r}
df_means <- function(df) {
  for(i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    print(col_mean)
  }
}
# colnames(df[i]) will return that column name as a string

df_means(df = mtcars)

# Update this function so that for each iteration it returns a phrase "The mean value of _____ is _____."

df_means <- function(df) {
  for(i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    print(paste("The mean value of", colnames(df[i]), "is", col_mean))
  }
}
```


Make a subset of only structural size variables from the penguins dataset for Chinstrap

```{r}
chinstrap_sizes <- penguins %>% 
  filter(species == "Chinstrap") %>% 
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)

df_means(df = chinstrap_sizes)
```

## Condition to check class of an input

```{r}
lunchtime <- function(tacos, price_per) {
  
  if(is.numeric(tacos) == FALSE) {
    stop("You must input a number of tacos.")
  }
  
  tot_price <- tacos * price_per
  return(tot_price)
}

lunchtime(tacos = 3, price_per = 2.50)
# lunchtime(tacos = "chicken", price_per = 2.50)
```

### Example: use values from a dataframe to calculate something based on user input.

```{r}
# Calculate the groundwater transport rate
gw_rate <- function(site) {
  
  gw_slope_depth <- data.frame(sitename = c("mountain", "prairie", "desert", "beach"),
                               depth = c(32, 42, 63, 2),
                               slope = c(11.2, 0.4, 0.8, 2.6))
  
  site_select <- dplyr::filter(gw_slope_depth, sitename == site)
  
  transport_rate <- 1.4 * site_select$slope + 3.6 * site_select$depth
  
  return(transport_rate)
}

gw_rate(site = "beach")
```

### Now using switch()

```{r}
gw_rate_switch <- function(site) {
  
  gw_slope_depth <- switch(site,
                           "mountain" = c(32, 11.2),
                           "prairie" = c(41, 0.4),
                           "desert" = c(63, 0.8),
                           "beach" = c(2, 2.6))
  transport_rate <- 1.4 * gw_slope_depth[2] + 3.6 * gw_slope_depth[1]
  
  return(transport_rate)
}

gw_rate_switch(site = "beach")
```

## Logistic growth equation

```{r}
logistic_growth <- function(K, N0, r, time) {
  Nt <-K / (1 + ((K-N0) / N0) * exp(-r * time))
  print(Nt)
}

logistic_growth(K = 5000, N0 = 200, r = 0.43, time = 6)

time_vec <- seq(from = 0, to = 35, by = 0.1)
bacterial_growth <- logistic_growth(K = 5000, N0 = 200, r = 0.43, time = time_vec)

bacterial_df <- data.frame(time_vec, bacterial_growth)

ggplot(data = bacterial_df, aes(x = time_vec, y = bacterial_growth)) +
  geom_point()

```


```{r}
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)
time_vec <- seq(from = 0, to = 35, by = 0.1)

# Create an empty matrix
out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

# N0 = 100, K = 6000
for(i in seq_along(r_seq)) {
  for(j in seq_along(time_vec)) {
    pop <- logistic_growth(K = 6000, N0 = 100, r = r_seq[i], time = time_vec[j])
    out_matrix[j, i] <- pop
  }
}

# Add time column
out_df <- data.frame(time = time_vec, out_matrix)

# Update column names
colnames(out_df) <- c("time", paste0("gr_", r_seq))

# Make a long form (tidy version) of data set
out_df_long <- out_df %>% 
  pivot_longer(-time, names_to = "growth_rate", values_to = "population")

# Make in ggplot a LINE GRAPH (geom_line()) for population (y-axis) over time (x-axis) with color depending on the growth rate

ggplot(data = out_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rate))
```







